Fig6

### 2020-04-21 analysis/figure-2.Rmd
now with candidate genes in wangle data, finally

```{r GO}

# all genes in this parental care dataset

geneids <- read_csv("../metadata/00_geneinfo.csv") %>%
  select(-X1) %>%
  rename("NCBI" = "entrezid",
         "gene" = "Name")

# From Ch 17 Evolution of parental care
curleychampagnegenes <- c("AVPR1A", "ADRA2A",    
                         "COMT", "CRH", "CRHBP", "CRHR1", "CRHR2", 
                          "DRD1", "DRD4","ESR1", "ESR2", #ERa ERbeta
                          "FOS", "HTR2C", "MEST", "NR3C1", #GR
                          "OPRM1", "PGR", "PRL", "PRLR",  "SLC6A4") #5HTT
curleychampagnegenes <- as.data.frame(curleychampagnegenes) %>%
  mutate(GO = "parentalcare", gene = curleychampagnegenes)  %>%
  select(GO, gene)
curleychampagnegenes

# Candidate GO terms 
GO_path <- "../metadata/goterms/"   # path to the data
GO_files <- dir(GO_path, pattern = "*.txt") # get file names
GO_pathfiles <- paste0(GO_path, GO_files)

GOgenesLong <- GO_pathfiles %>%
  setNames(nm = .) %>% 
  map_df(~read_table(.x, col_types = cols(), col_names = FALSE), .id = "file_name") %>% 
  mutate(GO = sapply(strsplit(as.character(file_name),'../metadata/goterms/'), "[", 2)) %>% 
  mutate(GO = sapply(strsplit(as.character(GO),'.txt'), "[", 1)) %>% 
  mutate(gene = sapply(strsplit(as.character(X1), "[\\\\]|[^[:print:]]" ), "[", 2)) %>% 
  select(GO, gene)  %>%
  filter(gene != "Symbol") %>%
  distinct(GO,gene)  %>%
  mutate(gene = toupper(gene)) %>%
  rbind(., curleychampagnegenes) %>%
  arrange(gene) %>%
  left_join(., geneids, by = "gene") %>%
  drop_na() 
GOgenesLong

GOgenesWide <- GOgenesLong %>% 
  pivot_wider(
    names_from = GO,
    values_from = gene) %>%
  mutate(numGOs = 4 - rowSums(is.na(.))) %>%
  arrange(desc(numGOs)) %>%
  select(-numGOs) %>%
  left_join(geneids, by = c("geneid","NCBI"))
GOgenesWide

# genes I can't find in dataset 
# NOS1 or OXTR or FOX1B or HTR5A
```

@@

```{r DEGs}

# summary DEG results from DESeq2
candidateDEGS <- read_csv("../results/suppletable1.csv") %>%
  filter(gene %in% candidategenes) %>%
  mutate(posneg = ifelse(lfc >= 0, "+", "-"),
         sex = recode(sex, "female" = "F", "male" = "M" ),
         tissue = recode(tissue, 
                         "hypothalamus" = "H",
                         "pituitary" = "P", "gonad" = "G")) %>%
  mutate(res = paste(sex, tissue, posneg, sep = "")) %>%
  select(gene, res, comparison)  %>%
  group_by(gene,  comparison) %>%
  summarize(res = str_c(res, collapse = " ")) %>%
  pivot_wider(names_from = comparison, values_from = res) %>%
  select(gene, bldg_lay, lay_inc.d3, inc.d3_inc.d9, inc.d9_inc.d17, hatch_n5, n5_n9)
candidateDEGS

>>>>

```{r GO}
source("../R/wrangledata.R")
```

@@

```{r DEGs}
# summary DEG results from DESeq2
candidateDEGS <- read_csv("../results/suppletable1.csv") %>%
  filter(gene %in% candidategenes) %>%
  mutate(posneg = ifelse(lfc >= 0, "+", "-"),
         sex = recode(sex, "female" = "F", "male" = "M" ),
         tissue = recode(tissue, 
                         "hypothalamus" = "H",
                         "pituitary" = "P", "gonad" = "G")) %>%
  mutate(res = paste(sex, tissue, posneg, sep = "")) %>%
  select(gene, res, comparison)  %>%
  group_by(gene,  comparison) %>%
  summarize(res = str_c(res, collapse = " ")) %>%
  pivot_wider(names_from = comparison, values_from = res) %>%
  select(gene, bldg_lay, lay_inc.d3, inc.d3_inc.d9, inc.d9_inc.d17, hatch_n5, n5_n9)
candidateDEGS

### 2020-04-21 analysis/figure-6.Rmd
now with candidate genes in wangle data, finally

```{r}
library(tidyverse)
library(ggtext)
library(cowplot)
library(ggpubr)
library(kableExtra)

@@

knitr::opts_chunk$set(echo = TRUE, message = F, fig.path = "../figures/")

candidategenesslim <- c("OXT", "AVP", "GNRH1", "GNRHR", 
                    "AR",  "CYP19A1", 
                     "AVPR1A", "AVPR1B", "AVPR2","VIP",
                  "DRD1", "DRD2", 
                  "PRL", "PRLR",  
                    "ESR1","ESR2", "LBH",

                    "FOS", "JUN", "EGR1", "BDNF"
                      )

candidategenes <- candidategenesslim

```

@@

```{r }
geneids <- read_csv("../metadata/00_geneinfo.csv")

vsd_path <- "../results/DEseq2/manip/"   # path to the data
vsd_files <- dir(vsd_path, pattern = "*vsd.csv") # get file names
vsd_pathfiles <- paste0(vsd_path, vsd_files)
vsd_files

@@

hypvsd <- getcandidatevsd(candidategenesslim, "hypothalamus", sexlevels)
pitvsd <- getcandidatevsd(candidategenesslim, "pituitary", sexlevels)
gonvsd <- getcandidatevsd(candidategenesslim, "gonad", sexlevels)
head(hypvsd)
head(pitvsd)
head(gonvsd)

@@

head(candidatevsd)
levels(candidatevsd$treatment)
```

@@

table1 <- allDEG %>%
  mutate(updown = ifelse(lfc > 0, "+", "-"))  %>%
  mutate(geneupdown = paste(gene, updown, sep = "")) %>%
  filter(gene %in% candidategenes) %>%
  arrange(geneupdown) %>%
  group_by(sex, tissue, comparison) %>%
  summarize(genes = str_c(geneupdown, collapse = " ")) %>%
  pivot_wider(names_from = comparison, values_from = genes ) %>%
  select(sex, tissue, contains("m."))  %>%
  arrange( tissue, sex)

(table1$inc.d3_m.inc.d3)
(table1$inc.d9_m.inc.d9)
(table1$inc.d17_m.inc.d17)
(table1$hatch_m.n2)
```

## Figs

```{r fig6, fig.height=7, fig.width=7}
makeboxplotsmanip <- function(df, whichgene, mysubtitle, whichsex, whichtimepoint){
  p <- df %>%
    filter(treatment %in% alllevels,
           gene %in% whichgene,
           sex %in% whichsex) %>%
    filter(treatment %in% whichtimepoint) %>%
    ggplot(aes(x = treatment, y = counts, fill = treatment, color = sex)) +
    geom_boxplot(outlier.shape = NA) + 
    geom_jitter(size = 0.5, width = 0.1) +
    facet_wrap(~gene, scales = "free_y", nrow = 1) +
    theme_B3() +
    scale_fill_manual(values = allcolors) +
    scale_color_manual(values = allcolors) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "none",
          plot.caption = element_text(face = "italic")) +
    labs(y = "gene expression" , x = "Parental Stages and Manipulations", subtitle = mysubtitle) +
    theme(strip.text = element_text(face = "italic"))
  return(p)
}

## hyp 
a2 <- makeboxplotsmanip(hypvsd, c("DRD1","EGR1","PRLR"), "Female hypothalamus", "female",
                  c("inc.d3", "m.inc.d3", "inc.d17", "m.inc.d17",  "hatch", "m.n2"))  +
  theme(axis.title.x = element_blank())
b2 <- makeboxplotsmanip(hypvsd, c("BDNF","ESR2", "FOS"), " ", "female",
                  c("inc.d3", "m.inc.d3", "inc.d17", "m.inc.d17"))  +
  theme(axis.title.x = element_blank())

c <- makeboxplotsmanip(hypvsd, c("AVP","OXT"), " ", "female", 
                       c("inc.d17", "m.inc.d17"))  +
  theme(axis.title = element_blank())
d <- makeboxplotsmanip(hypvsd, c("AVPR1A","CYP19A1","GNRH1"), " ", "female",
                   c("hatch", "m.n2"))  +
  theme(axis.title = element_blank())
e <- makeboxplotsmanip(hypvsd, c("JUN"), " ", "female", 
                 c("hatch", "m.n2"))  +
  theme(axis.title = element_blank()) + labs(subtitle = " ")

f1 <- makeboxplotsmanip(hypvsd, c("DRD1", "EGR1", "PRLR"), "Male hypothalamus", "male",
                  c( "inc.d17", "m.inc.d17", "hatch", "m.n2"))  +
  theme(axis.title.x = element_blank())

f2 <- makeboxplotsmanip(hypvsd, c( "PRL"), " ", "male",
                  c( "inc.d17", "m.inc.d17", "hatch", "m.n2"))  +
  theme(axis.title = element_blank())

h <- makeboxplotsmanip(hypvsd, c("AR", "AVP", "BDNF", "ESR1", "LBH"), " ", "male",
                  c( "inc.d17", "m.inc.d17"))  + labs(caption = " ")

i <- makeboxplotsmanip(hypvsd, c("GNRH1"), " ", "male",
                  c("hatch", "m.n2"))  +
  theme(axis.title = element_blank()) +
  labs(x = " ",
       caption = "Candidate genes that were not differentially expressed following removal: AVPR1B, AVPR2, ESR1, GNRHR, VIP")

ab2 <- plot_grid(a2,c, rel_widths = c(18,5), labels = "auto", label_size = 8)
cde <- plot_grid(b2,d,e, rel_widths = c(12,6,2), nrow =  1, align = "h",  
                 labels = c("c", "d", " "), label_size = 8)
f12 <- plot_grid(f1,f2, rel_widths = c(3,1), labels = c("e",  " "), label_size = 8)
hi <- plot_grid(h,i, ncol = 2, rel_widths = c(10,2),labels = c("f",  "g"), label_size = 8, align = "h")

plot_grid(ab2, cde, f12,hi, nrow = 4, rel_heights = c(1,1,1,1.25))

```

```{r writefiles}
write.csv(candidatevsd, "../../musicalgenes/data/candidatecounts.csv")
```

>>>>

```{r setup}
library(tidyverse)
library(ggtext)
library(cowplot)
library(ggpubr)
library(kableExtra)

@@

knitr::opts_chunk$set(echo = TRUE, message = F, fig.path = "../figures/")
```

## Candidate Genes (from literature and relevant GO terms)

```{r GO}
source("../R/wrangledata.R")
```

@@

```{r vsd}
vsd_path <- "../results/DEseq2/manip/"   # path to the data
vsd_files <- dir(vsd_path, pattern = "*vsd.csv") # get file names
vsd_pathfiles <- paste0(vsd_path, vsd_files)
vsd_files

@@

hypvsd <- getcandidatevsd(candidategenes, "hypothalamus", sexlevels)
pitvsd <- getcandidatevsd(candidategenes, "pituitary", sexlevels)
gonvsd <- getcandidatevsd(candidategenes, "gonad", sexlevels)

@@

head(candidatevsd)

levels(candidatevsd$treatment)
```

@@

candidateDEGS <- allDEG %>%
  filter(gene %in% candidategenes) %>%
  mutate(posneg = ifelse(lfc >= 0, "+", "-"),
         sex = recode(sex, "female" = "F", "male" = "M" ),
         tissue = recode(tissue, 
                         "hypothalamus" = "H",
                         "pituitary" = "P", "gonad" = "G")) %>%
  mutate(res = paste(sex, tissue, posneg, sep = "")) %>%
  select(gene, res, comparison)  %>%
  group_by(gene,  comparison) %>%
  summarize(res = str_c(res, collapse = " ")) %>%
  pivot_wider(names_from = comparison, values_from = res) %>%
  select(gene, contains("m.")) 
candidateDEGS

table2 <- left_join(candidateDEGS, GOgenesWide) %>%
  select(gene, m.inc.d17_prolong:m.inc.d9_m.inc.d17, parentalcare, parentalbehavior, NCBI) %>%
  mutate(parentalcare = if_else(is.na(parentalcare), " ", "X"),
         parentalbehavior = if_else(is.na(parentalbehavior), " ", "X")) %>%
  rename("Literature" = "parentalcare", "GO" =  "parentalbehavior")
table2$numDEGs <- rowSums(is.na(table2)) # count NAs to know how many are NS
table2 <- table2 %>% 
  mutate(sig = ifelse(numDEGs == 6, "NS", "DEG")) %>% 
  arrange(sig, gene)  %>%  select(-sig, -numDEGs)
table2[is.na(table2)] <- " " # replace NA with blank space so it's pretty
table2
```

```{r writefiles}
write.csv(table2, "../results/table2.csv")

# for victoria
prlmanip <- candidatevsd %>%
  filter(gene == "PRL", 
         treatment %in% c("extend", "m.inc.d8"))
write.csv(prlmanip, "../results/prlmanip.csv")
```
